---
title: "crussol.clementine"
author: "Clémentine Crussol"
output:
  html_document:
    df_print: paged
    code_folding: show
    toc: false
    number_sections: true
---
# Impaired quality of life after intensive care

# Scientific problem 
## Context

Advances in intensive care medicine have significantly reduced ICU mortality over recent decades, leading to a growing population of ICU survivors with persistent impairments after discharge (1, 2). These long-term consequences, commonly grouped under the term Post-Intensive Care Syndrome (PICS), may substantially alter patients’ health-related quality of life (HRQoL) (3).

According to the World Health Organization (WHO), quality of life refers to an individual’s perception of their position in life within the cultural and social context in which they live (4). In the medical field, HRQoL is considered a multidimensional concept reflecting the influence of health status and healthcare on physical, psychological, and social dimensions of health (5). Previous studies have shown that ICU survivors often report lower HRQoL scores than the general population, with only partial recovery over time (6, 7).

This project aims to describe the evolution of HRQoL at 3 and 12 months after ICU discharge using the SF-36 questionnaire.

## Aim

The aim of this study is to evaluate the quality of life of ICU survivors at 3 and 12 months after discharge, using the SF-36 questionnaire. Specifically, it seeks to assess changes over time in the different dimensions of quality of life, including physical and mental components, in order to better understand recovery trajectories following critical illness.

## Method

All ICU survivors were contacted by mail 3 and 12 month after hospital discharge and were asked to complete the SF-36 questionnaire.

### Outcome measure (8, 9)

Quality of life was assesed using the SF-36 questionnaire. This questionnaire evaluate multiple dimensions of quality of life including : Physical Functionning (PF), Role Physical (RP), Bodily Pain (BP), General Health (GH), Vitality (VT), Social Functioning (SF), Role Emotional (RE), Mental Health (MH). 
Then, this dimensions are grouped into two components, the Physical Component Summary (PCS), including PF, RP, BP and GH, and the Mental Component Summary (MCS), including VT, SF, RE and MH.

SF-36 domains were scored according to standard procedures. Scores of each dimension were first computed by summing the values of all items belonging to that dimension and were then linearly transformed to a 0–100 scale. Higher scores indicate better heath status.

The Physical Component Summary (PCS) and Mental Component Summary (MCS) were then calculated using the dimensions scores. Then they were standardized unsing French reference population norms, with a mean of 50 and a standard deviation of 10.

Detailed information on SF-36 scoring procedures, including item mapping, score transformation, and computation of PCS and MCS, is provided in the `SF-36 appendix.md`.

### Participants

479 patients who survived ICU and responded to the SF-36 at both 3 and 12 months were included in the study.

# RStudio project
## Aim

The aim of this project is to assess SF-36 dimension scores at 3 and 12 months after ICU discharge. In addition, a radar chart will be generated to visualize changes in quality of life dimensions over time.

## Data organization
File: `bazRSF.csv`  
Colums:  
`ID`: Patient ID, corresponding to their number of inclusion  
`SF_3m_Q[y]`: SF-36 score at 3 months for question y (where y is the question number e.g. 1, 2, 3a, ...)   
`SF_12m_Q[y]`: SF-36 score at 12 months for question y (where y is the question number e.g. 1, 2, 3a, ...)  

## Script organization
### Recoding of SF-36 item scores
**Aim**: To recode the SF-36 item responses so that higher scores correspond to a better health status accross all items.  
**Input**: `bazRSF.csv` in `data` folder  
**Output**: A set of SF-36 item scores in which higher values indicate better perceived health status.

### Calculation and transformation of dimension scores at 3 and 12 months
**Aim**: To compute dimension scores and transform them onto a 0-100 scale for comparability and interpretation  
**Input**: `bazRSF.csv` in `data` folder  
**Calculation**:   
Before computing dimension scores, we perform mean imputation at the dimension level. If at least 50% of items within a dimension are available, missing values are replaced by the mean of the observed items in that dimension. If fewer than 50% of items are present, the dimension score is left as `NA`.  

Then, raw scores for each dimension are computed by summing the values of all items belonging to that dimension.  
Scores are then linearly transformed to a 0–100 scale using the following formula:  

$$
\text{Transformed score} = \frac{\text{Observed score} - \text{Minimum possible score}}{\text{Maximum possible score} - \text{Minimum possible score}} \times 100
$$

In this transformed scale, 0 represents the worst possible health status and 100 represents the best possible health status.  
**Output**: A standardized score for each dimension, ranging between 0 and 100  

### Visualization of the evolution of dimension scores between 3 and 12 months with a radar plot
**Aim**: To visually compare dimension scores between 3 and 12 months, and assess if changes over time is statistically significant, using a radar plot  
**Input**: `bazRSF.csv` in `data` folder  
**Output**: A radar plot comparing dimension scores at 3 and 12 months with p-values indicating whether changes over time reflect a significant improvement or not  

---

## Code

### Setup

These lines load the necessary libraries for data manipulation and visualization.
```{r}
# If you don't have the packages installed, you can install it using the following commands:
# install.packages("fmsb")

# Then, load the library:

library(fmsb)

```

```{r}
# Load the data
bazR <- read.csv("data/bazRSF.csv", header = TRUE, sep = ";")

```

### Recoding of SF-36 item scores
Item responses are recoded so that higher values reflect better health status across all SF-36 dimensions, ensuring comparability between items and correct interpretation of scores. The items that need to be recoded are Q1, Q2, Q6, Q7, Q8, Q9a, Q9d, Q9e, Q9h, Q11b and Q11d at both 3 and 12 months.

```{r}
# 3 months
# Items requiring reverse coding are recoded using a linear transformation (new score = max + 1 - original score) in order to ensure that higher scores reflect better health status.

## Define functions to reverse code items with 5-point and 6-point scales
reverse6 <- function(x) 6 - x
reverse7 <- function(x) 7 - x

## Items with 5-point scale (Q1, Q2, Q6, Q11b, Q11d)

items_6 <- c("SF_3m_Q1","SF_3m_Q2","SF_3m_Q6","SF_3m_Q8",
             "SF_3m_Q11b","SF_3m_Q11d",
             "SF_12m_Q1","SF_12m_Q2","SF_12m_Q6",
             "SF_12m_Q8","SF_12m_Q11b","SF_12m_Q11d")

## Items with 6-point scale (Q7, Q9a, Q9d, Q9e, Q9h)

items_7 <- c("SF_3m_Q7","SF_3m_Q9a","SF_3m_Q9d","SF_3m_Q9e","SF_3m_Q9h",
             "SF_12m_Q7","SF_12m_Q9a","SF_12m_Q9d","SF_12m_Q9e","SF_12m_Q9h")

## Apply reverse coding to the specified items

bazR[items_6] <- lapply(bazR[items_6], reverse6)
bazR[items_7] <- lapply(bazR[items_7], reverse7)

```

### Calculation and transformation of dimension scores at 3 and 12 months
#### Imputation of missing data
Missing item values were imputed at the dimension level when at least 50% of items were available: in this case, missing values were replaced by the mean of observed items within the same dimension. If less than 50% of items were present, the dimension score was not computed (NA).

```{r}
# Each dimension is defined as a set of questionnaire items grouped according to SF-36 scoring guidelines
# Group questionnaire items into their respective health dimension according to the standard SF-36 classification


items_sf36_3m <- list(
  PF = c("SF_3m_Q3a", "SF_3m_Q3b","SF_3m_Q3c","SF_3m_Q3d", "SF_3m_Q3e", "SF_3m_Q3f","SF_3m_Q3g",
         "SF_3m_Q3h","SF_3m_Q3i","SF_3m_Q3j"),
  RP = c("SF_3m_Q4a", "SF_3m_Q4b","SF_3m_Q4c","SF_3m_Q4d"),
  BP = c("SF_3m_Q7","SF_3m_Q8"),
  GH = c("SF_3m_Q1","SF_3m_Q11a","SF_3m_Q11b", "SF_3m_Q11c","SF_3m_Q11d"),
  VT = c("SF_3m_Q9a", "SF_3m_Q9e","SF_3m_Q9g","SF_3m_Q9i"),
  SF = c("SF_3m_Q6", "SF_3m_Q10"),
  RE = c("SF_3m_Q5a","SF_3m_Q5b","SF_3m_Q5c"),
  MH = c("SF_3m_Q9b","SF_3m_Q9c","SF_3m_Q9d","SF_3m_Q9f","SF_3m_Q9h")
)

# Function to impute missing SF-36 item values at the dimension level

impute_sf36 <- function(df, items_list) {
  df_out <- df # copy the dataset to not modify directly the original data
  
  for (dim in names(items_list)) {
    vars <- items_list[[dim]]
    
# Count number of non-missing items per participant for each dimension
    present <- rowSums(!is.na(df[vars]))
    
# SF-36 rule: at least 50% of items must be available to compute the dimension score
    enough <- present >= length(vars) / 2
    
# Compute row-wise mean of available items for imputation
    row_means <- rowMeans(df[vars], na.rm = TRUE)
    
# Impute missing values using dimension mean when rule is satisfied
    for (v in vars) {
      df_out[[v]][is.na(df[[v]]) & enough] <- row_means[is.na(df[[v]]) & enough]
    }
    
# If insufficient data, set entire dimension to NA
    df_out[!enough, vars] <- NA
  }
  
  return(df_out)
}

# Apply imputation function to dataset
bazR <- impute_sf36(bazR, items_sf36_3m)
```

#### Calculation and transformation of dimension scores at 3 months
Raw scores for each dimension were first computed by summing the values of all items belonging to that dimension.
Scores were then linearly transformed to a 0–100 scale using the following formula:

Transformed score = ((Observed score - Minimum possible score) / (Maximum possible score - Minimum possible score)) × 100

In this transformed scale, 0 represents the worst possible health status and 100 represents the best possible health status.

```{r}
# Compute dimension PF
aPF_3m <- (bazR$SF_3m_Q3a + bazR$SF_3m_Q3b + bazR$SF_3m_Q3c + bazR$SF_3m_Q3d + bazR$SF_3m_Q3e + bazR$SF_3m_Q3f
            + bazR$SF_3m_Q3g + bazR$SF_3m_Q3h + bazR$SF_3m_Q3i + bazR$SF_3m_Q3j)

# Transform dimension PF to a 0-100 scale
bazR$PFtransformed_3m <- ((aPF_3m-10)/20)*100 


# Do the same for all others dimensions 
# Dimension RP
aRP_3m <- (bazR$SF_3m_Q4a + bazR$SF_3m_Q4b + bazR$SF_3m_Q4c + bazR$SF_3m_Q4d)

bazR$RPtransformed_3m<-((aRP_3m-4)/4)*100

# Dimension BP
aBP_3m <- (bazR$SF_3m_Q7 + bazR$SF_3m_Q8)

bazR$BPtransformed_3m <- ((aBP_3m-2)/10)*100

# Dimension GH
aGH_3m <- (bazR$SF_3m_Q1 + bazR$SF_3m_Q11a + bazR$SF_3m_Q11b + bazR$SF_3m_Q11c + bazR$SF_3m_Q11d)

bazR$GHtransformed_3m <- ((aGH_3m-5)/20)*100

# Dimension VT
aVT_3m <- (bazR$SF_3m_Q9a + bazR$SF_3m_Q9e + bazR$SF_3m_Q9g + bazR$SF_3m_Q9i)

bazR$VTtransformed_3m <- ((aVT_3m-4)/20)*100

# Dimension SF
aSF_3m <- (bazR$SF_3m_Q6 + bazR$SF_3m_Q10)

bazR$SFtransformed_3m <- ((aSF_3m-2)/8)*100

#calcul dimension RE
aRE_3m <- (bazR$SF_3m_Q5a + bazR$SF_3m_Q5b + bazR$SF_3m_Q5c)

bazR$REtransformed_3m <- ((aRE_3m-3)/3)*100

# Dimension MH
aMH_3m <- (bazR$SF_3m_Q9b + bazR$SF_3m_Q9c + bazR$SF_3m_Q9d + bazR$SF_3m_Q9f + bazR$SF_3m_Q9h)

bazR$MHtransformed_3m <- ((aMH_3m-5)/25)*100

# Dimension HT
aHT_3m <- bazR$SF_3m_Q2

bazR$HTtransformed_3m <- ((aHT_3m-1)/4)*100

```


#### Imputation of missing data at 12 months
Missing item values were imputed using the same method and function as for 3 months, with the same rule of at least 50% of items available to compute the dimension score.

```{r}

items_sf36_12m <- list(
  PF = c("SF_12m_Q3a", "SF_12m_Q3b","SF_12m_Q3c","SF_12m_Q3d", "SF_12m_Q3e", "SF_12m_Q3f",
         "SF_12m_Q3g", "SF_12m_Q3h","SF_12m_Q3i","SF_12m_Q3j"),
  RP = c("SF_12m_Q4a", "SF_12m_Q4b", "SF_12m_Q4c","SF_12m_Q4d"),
  BP = c("SF_12m_Q7","SF_12m_Q8"),
  GH = c("SF_12m_Q1","SF_12m_Q11a","SF_12m_Q11b", "SF_12m_Q11c","SF_12m_Q11d"),
  VT = c("SF_12m_Q9a", "SF_12m_Q9e","SF_12m_Q9g","SF_12m_Q9i"),
  SF = c("SF_12m_Q6", "SF_12m_Q10"),
  RE = c("SF_12m_Q5a","SF_12m_Q5b","SF_12m_Q5c"),
  MH = c("SF_12m_Q9b","SF_12m_Q9c","SF_12m_Q9d","SF_12m_Q9f","SF_12m_Q9h")
)


bazR <- impute_sf36(bazR, items_sf36_12m)

```

#### Calculation and transformation of dimension scores at 3 months
Scores for each dimension were computed using the same method as for 3 months, by summing the values of all items belonging to that dimension and then linearly transforming them to a 0–100 scale using the same formula as for 3 months.

```{r}
# Dimension PF
aPF_12m <- (bazR$SF_12m_Q3a + bazR$SF_12m_Q3b + bazR$SF_12m_Q3c + bazR$SF_12m_Q3d + bazR$SF_12m_Q3e 
            + bazR$SF_12m_Q3f + bazR$SF_12m_Q3g + bazR$SF_12m_Q3h + bazR$SF_12m_Q3i + bazR$SF_12m_Q3j)

bazR$PFtransformed_12m<-((aPF_12m-10)/20)*100 

# Dimension RP
aRP_12m <- (bazR$SF_12m_Q4a + bazR$SF_12m_Q4b + bazR$SF_12m_Q4c + bazR$SF_12m_Q4d)

bazR$RPtransformed_12m <- ((aRP_12m-4)/4)*100

# Dimension BP
aBP_12m <- (bazR$SF_12m_Q7 + bazR$SF_12m_Q8)

bazR$BPtransformed_12m <- ((aBP_12m-2)/10)*100

# Dimension GH
aGH_12m <- (bazR$SF_12m_Q1 + bazR$SF_12m_Q11a + bazR$SF_12m_Q11b + bazR$SF_12m_Q11c + bazR$SF_12m_Q11d)

bazR$GHtransformed_12m <- ((aGH_12m-5)/20)*100

# Dimension VT
aVT_12m <- (bazR$SF_12m_Q9a + bazR$SF_12m_Q9e + bazR$SF_12m_Q9g + bazR$SF_12m_Q9i)

bazR$VTtransformed_12m <- ((aVT_12m-4)/20)*100

# Dimension SF
aSF_12m <- (bazR$SF_12m_Q6 + bazR$SF_12m_Q10)

bazR$SFtransformed_12m <- ((aSF_12m-2)/8)*100

# Dimension RE
aRE_12m <- (bazR$SF_12m_Q5a + bazR$SF_12m_Q5b + bazR$SF_12m_Q5c)

bazR$REtransformed_12m <- ((aRE_12m-3)/3)*100

# Dimension MH
aMH_12m <- (bazR$SF_12m_Q9b + bazR$SF_12m_Q9c + bazR$SF_12m_Q9d + bazR$SF_12m_Q9f + bazR$SF_12m_Q9h)

bazR$MHtransformed_12m <- ((aMH_12m-5)/25)*100

# Dimension HT
aHT_12m <- bazR$SF_12m_Q2

bazR$HTtransformed_12m <- ((aHT_12m-1)/4)*100

```

### Visualization of the evolution of dimension scores between 3 and 12 months with a radar plot

```{r}
# Compare SF-36 scores between 3 and 12 months and store values for the construction of radar plot. And perform Wilcoxon test to assess if changes over time are statistically significant, with a significance threshold of p<0.05.

# Dimension PF
summary(bazR$PFtransformed_3m)
summary(bazR$PFtransformed_12m)
radarPF=c(100,0,54.67,59.29) #max 100, min 0, mean 3m, mean 12m
wilcox.test(bazR$PFtransformed_3m,bazR$PFtransformed_12m)

# Dimension RP
summary(bazR$RPtransformed_3m)
summary(bazR$RPtransformed_12m)
radarRP=c(100,0,35.19,54.63)
wilcox.test(bazR$RPtransformed_3m,bazR$RPtransformed_12m)

# Dimension BP
summary(bazR$BPtransformed_3m)
summary(bazR$BPtransformed_12m)
radarBP=c(100,0,36.33,51.67)
wilcox.test(bazR$BPtransformed_3m,bazR$BPtransformed_12m)

# Dimension GH
summary(bazR$GHtransformed_3m)
summary(bazR$GHtransformed_12m)
radarHG=c(100,0,40.7,58.83)
wilcox.test(bazR$GHtransformed_3m,bazR$GHtransformed_12m)

# Dimension VT
summary(bazR$VTtransformed_3m)
summary(bazR$VTtransformed_12m)
radarVT=c(100,0,42.22,50.50)
wilcox.test(bazR$VTtransformed_3m,bazR$VTtransformed_12m)

# Dimension SF
summary(bazR$SFtransformed_3m)
summary(bazR$SFtransformed_12m)
radarSF=c(100,0,51.67,67.92)
wilcox.test(bazR$SFtransformed_3m,bazR$SFtransformed_12m)

# Dimension RE
summary(bazR$REtransformed_3m)
summary(bazR$REtransformed_12m)
radarRE=c(100,0,51.28,62.96)
wilcox.test(bazR$REtransformed_3m,bazR$REtransformed_12m)

# Dimension MH
summary(bazR$MHtransformed_3m)
summary(bazR$MHtransformed_12m)
radarMH=c(100,0,58.80,61.47)
wilcox.test(bazR$MHtransformed_3m,bazR$MHtransformed_12m)

# Create a matrix for the radar plot with the computed scores for each dimension at 3 and 12 months, along with the maximum and minimum values (0-100)
data<-cbind(radarPF,radarRP,radarBP,radarHG,radarVT,radarSF,radarRE,radarMH) #fait tableau avec ensemble dimensions

# Define labels for each dimension that include the p-values from the Wilcoxon tests

labels = c("PF \n p=0.5","RP\n p=0.1","BP\n p=0.2","GH\n p<0.0001",
           "VT\n p=0.01","SF\n  p=0.01","RE\n p=0.3","MH\n p=0.8")

# Rename columns of the matrix with the defined labels
colnames(data)<-labels

# Convert the matrix to a data frame for use with the radar chart function
datx = as.data.frame(data)

coul <- c("red", "blue") # Define colors for the radar chart: red for 3 months and blue for 12 months

# Save radar plot as a PNG file
png("./results/radar_plot.png",width = 2400,height = 1800,res = 300)

# Create the radar chart using the fmsb package, with specified parameters for axis type, segments, colors, line width and type, grid color and type, and axis label color
radarchart(datx,
           axistype = 2, # graduation of axes
           seg = 5, # number of segments in the radar chart
           vlcex = 0.7, # size of labels for each dimension
           pcol = coul,# colors for the lines representing 3 and 12 months
           plty=1, # style of lines
           cglcol="gray71", # color of the grid lines
           cglty=1, # style of grid lines
           axislabcol="grey", # color of axis labels

)


legend(1.2,0.8,# position of the legend on the plot
       legend=c("3 month","12 month"), 
       pch=c(16,16), # symbols
       col=c("red","blue"), # colors for the symbols
       lty=c(1,1), # line types
)
dev.off()
```

## Export the data in the results folder

```{r}
write.csv(bazR, "results/bazP_scores.csv", row.names = FALSE)
```

