---
title: "crussol.clementine"
author: "Clémentine Crussol"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: false
    number_sections: true
---

##  Import the data

```{r}
bazR <- read.csv("data/bazRSF_reseau.csv", header = TRUE, sep = ";")
```


## Calculation of dimension scores at 3 months
### Imputation of missing data

```{r}
items_sf36 <- list(
  PF = c("SF_3mQ3a", "SF_3mQ3b","SF_3mQ3c","SF_3mQ3d", "SF_3mQ3e", "SF_3mQ3f","SF_3mQ3g", "SF_3mQ3h","SF_3mQ3i","SF_3mQ3j"),
  RP = c("SF_3mQ4a", "SF_3mQ4b","SF_3mQ4c","SF_3mQ4d"),
  BP = c("SF_3mQ7","SF_3mQ8"),
  GH = c("SF_3mQ1","SF_3mQ11a","SF_3mQ11b", "SF_3mQ11c","SF_3mQ11d"),
  VT = c("SF_3mQ9a", "SF_3mQ9e","SF_3mQ9g","SF_3mQ9i"),
  SF = c("SF_3mQ6", "SF_3mQ10"),
  RE = c("SF_3mQ5a","SF_3mQ5b","SF_3mQ5c"),
  MH = c("SF_3mQ9b","SF_3mQ9c","SF_3mQ9d","SF_3mQ9f","SF_3mQ9h")
)

impute_sf36 <- function(df, items_list) {
  df_out <- df
  
  for (dim in names(items_list)) {
    vars <- items_list[[dim]]
    
    # nombre d'items présents par patient
    present <- rowSums(!is.na(df[vars]))
    
    # règle SF-36 : au moins 50% des items présents
    enough <- present >= length(vars) / 2
    
    # moyenne des items présents
    row_means <- rowMeans(df[vars], na.rm = TRUE)
    
    # imputation des NA si règle respectée
    for (v in vars) {
      df_out[[v]][is.na(df[[v]]) & enough] <- row_means[is.na(df[[v]]) & enough]
    }
    
    # si < 50% → toute la dimension devient NA
    df_out[!enough, vars] <- NA
  }
  
  return(df_out)
}
bazR <- impute_sf36(bazR, items_sf36)
```

### Calculation and transformation of dimension scores at 3 months
Raw scores for each dimension were first computed by summing the values of all items belonging to that dimension.
Scores were then linearly transformed to a 0–100 scale using the following formula:

Transformed score = ((Observed score - Minimum possible score) / (Maximum possible score - Minimum possible score)) × 100

In this transformed scale, 0 represents the worst possible health status and 100 represents the best possible health status.

```{r}
# Dimension PF
aPF_3m <- (bazR$SF_3mQ3a + bazR$SF_3mQ3b + bazR$SF_3mQ3c + bazR$SF_3mQ3d + bazR$SF_3mQ3e + bazR$SF_3mQ3f
            + bazR$SF_3mQ3g + bazR$SF_3mQ3h + bazR$SF_3mQ3i + bazR$SF_3mQ3j)

bazR$PFtransformed_3m <- ((aPF_3m-10)/20)*100 


# Dimension RP  
aRP_3m <- (bazR$SF_3mQ4a + bazR$SF_3mQ4b + bazR$SF_3mQ4c + bazR$SF_3mQ4d)

bazR$RPtransformed_3m<-((aRP_3m-4)/4)*100

# Dimension BP
aBP_3m <- (bazR$SF_3mQ7 + bazR$SF_3mQ8)

bazR$BPtransformed_3m <- ((aBP_3m-2)/10)*100

# Dimension GH
aGH_3m <- (bazR$SF_3mQ1 + bazR$SF_3mQ11a + bazR$SF_3mQ11b + bazR$SF_3mQ11c + bazR$SF_3mQ11d)

bazR$GHtransformed_3m <- ((aGH_3m-5)/20)*100

# Dimension VT
aVT_3m <- (bazR$SF_3mQ9a + bazR$SF_3mQ9e + bazR$SF_3mQ9g + bazR$SF_3mQ9i)

bazR$VTtransformed_3m <- ((aVT_3m-4)/20)*100

# Dimension SF
aSF_3m <- (bazR$SF_3mQ6 + bazR$SF_3mQ10)

bazR$SFtransformed_3m <- ((aSF_3m-2)/8)*100

#calcul dimension RE
aRE_3m <- (bazR$SF_3mQ5a + bazR$SF_3mQ5b + bazR$SF_3mQ5c)

bazR$REtransformed_3m <- ((aRE_3m-3)/3)*100

# Dimension MH
aMH_3m <- (bazR$SF_3mQ9b + bazR$SF_3mQ9c + bazR$SF_3mQ9d + bazR$SF_3mQ9f + bazR$SF_3mQ9h)

bazR$MHtransformed_3m <- ((aMH_3m-5)/25)*100

# Dimension HT
aHT_3m <- bazR$SF_3mQ2

bazR$HTtransformed_3m <- ((aHT_3m-1)/4)*100

```


## Calculation of dimension scores at 12 months
### Imputation of missing data

```{r}

items_sf36 <- list(
  PF = c("SF_12mQ3a", "SF_12mQ3b","SF_12mQ3c","SF_12mQ3d", "SF_12mQ3e", "SF_12mQ3f",
         "SF_12mQ3g", "SF_12mQ3h","SF_12mQ3i","SF_12mQ3j"),
  RP = c("SF_12mQ4a", "SF_12mQ4b", "SF_12mQ4c","SF_12mQ4d"),
  BP = c("SF_12mQ7","SF_12mQ8"),
  GH = c("SF_12mQ1","SF_12mQ11a","SF_12mQ11b", "SF_12mQ11c","SF_12mQ11d"),
  VT = c("SF_12mQ9a", "SF_12mQ9e","SF_12mQ9g","SF_12mQ9i"),
  SF = c("SF_12mQ6", "SF_12mQ10"),
  RE = c("SF_12mQ5a","SF_12mQ5b","SF_12mQ5c"),
  MH = c("SF_12mQ9b","SF_12mQ9c","SF_12mQ9d","SF_12mQ9f","SF_12mQ9h")
)

impute_sf36 <- function(df, items_list) {
  df_out <- df
  
  for (dim in names(items_list)) {
    vars <- items_list[[dim]]
    
    # nombre d'items présents par patient
    present <- rowSums(!is.na(df[vars]))
    
    # règle SF-36 : au moins 50% des items présents
    enough <- present >= length(vars) / 2
    
    # moyenne des items présents
    row_means <- rowMeans(df[vars], na.rm = TRUE)
    
    # imputation des NA si règle respectée
    for (v in vars) {
      df_out[[v]][is.na(df[[v]]) & enough] <- row_means[is.na(df[[v]]) & enough]
    }
    
    # si < 50% → toute la dimension devient NA
    df_out[!enough, vars] <- NA
  }
  
  return(df_out)
  
}
bazR <- impute_sf36(bazR, items_sf36)

```

### Calculation and transformation of dimension scores at 3 months
Raw scores for each dimension were first computed by summing the values of all items belonging to that dimension.
Scores were then linearly transformed to a 0–100 scale using the following formula:

Transformed score = ((Observed score - Minimum possible score) / (Maximum possible score - Minimum possible score)) × 100

In this transformed scale, 0 represents the worst possible health status and 100 represents the best possible health status.

```{r}
# Dimension PF
aPF_12m <- (bazR$SF_12mQ3a + bazR$SF_12mQ3b + bazR$SF_12mQ3c + bazR$SF_12mQ3d + bazR$SF_12mQ3e 
            + bazR$SF_12mQ3f + bazR$SF_12mQ3g + bazR$SF_12mQ3h + bazR$SF_12mQ3i + bazR$SF_12mQ3j)

bazR$PFtransformed_12m<-((aPF_12m-10)/20)*100 

# Dimension RP
aRP_12m <- (bazR$SF_12mQ4a + bazR$SF_12mQ4b + bazR$SF_12mQ4c + bazR$SF_12mQ4d)

bazR$RPtransformed_12m <- ((aRP_12m-4)/4)*100

# Dimension BP
aBP_12m <- (bazR$SF_12mQ7 + bazR$SF_12mQ8)

bazR$BPtransformed_12m <- ((aBP_12m-2)/10)*100

# Dimension GH
aGH_12m <- (bazR$SF_12mQ1 + bazR$SF_12mQ11a + bazR$SF_12mQ11b + bazR$SF_12mQ11c + bazR$SF_12mQ11d)

bazR$GHtransformed_12m <- ((aGH_12m-5)/20)*100

# Dimension VT
aVT_12m <- (bazR$SF_12mQ9a + bazR$SF_12mQ9e + bazR$SF_12mQ9g + bazR$SF_12mQ9i)

bazR$VTtransformed_12m <- ((aVT_12m-4)/20)*100

# Dimension SF
aSF_12m <- (bazR$SF_12mQ6 + bazR$SF_12mQ10)

bazR$SFtransformed_12m <- ((aSF_12m-2)/8)*100

# Dimension RE
aRE_12m <- (bazR$SF_12mQ5a + bazR$SF_12mQ5b + bazR$SF_12mQ5c)

bazR$REtransformed_12m <- ((aRE_12m-3)/3)*100

# Dimension MH
aMH_12m <- (bazR$SF_12mQ9b + bazR$SF_12mQ9c + bazR$SF_12mQ9d + bazR$SF_12mQ9f + bazR$SF_12mQ9h)

bazR$MHtransformed_12m <- ((aMH_12m-5)/25)*100

# Dimension HT
aHT_12m <- bazR$SF_12mQ2

bazR$HTtransformed_12m <- ((aHT_12m-1)/4)*100

```

## Statistical comparison of dimension scores at 3 and 12 months and construction of a radar plot

```{r}

summary(bazR$PFtransformed_3m)
summary(bazR$PFtransformed_12m)
radarPF=c(100,0,57.94,61.19) #max 100, min 0, PF3mois=58.19, PF1an=62.27
wilcox.test(bazR$PFtransformed_3m,bazR$PFtransformed_12m)


summary(bazR$RPtransformed_3m)
summary(bazR$RPtransformed_12m)
radarRP=c(100,0,31.6,47.03)
wilcox.test(bazR$RPtransformed_3m,bazR$RPtransformed_12m)


summary(bazR$BPtransformed_3m)
summary(bazR$BPtransformed_12m)
radarBP=c(100,0,46.64,48.16)
wilcox.test(bazR$BPtransformed_3m,bazR$BPtransformed_12m)


summary(bazR$GHtransformed_3m)
summary(bazR$GHtransformed_12m)
radarHG=c(100,0,50.42,52.3)
wilcox.test(bazR$GHtransformed_3m,bazR$GHtransformed_12m)


summary(bazR$VTtransformed_3m)
summary(bazR$VTtransformed_12m)
radarVT=c(100,0,47.53,51.22)
wilcox.test(bazR$VTtransformed_3m,bazR$VTtransformed_12m)


summary(bazR$SFtransformed_3m)
summary(bazR$SFtransformed_12m)
radarSF=c(100,0,61.01,67.09)
wilcox.test(bazR$SFtransformed_3m,bazR$SFtransformed_12m)


summary(bazR$REtransformed_3m)
summary(bazR$REtransformed_12m)
radarRE=c(100,0,46.19,56.06)
wilcox.test(bazR$REtransformed_3m,bazR$REtransformed_12m)


summary(bazR$MHtransformed_3m)
summary(bazR$MHtransformed_12m)
radarMH=c(100,0,61.35,62.45)
wilcox.test(bazR$MHtransformed_3m,bazR$MHtransformed_12m)

png("./results/radar_plot.png",width = 1800,height = 1800,res = 300)

data<-cbind(radarPF,radarRP,radarBP,radarHG,radarVT,radarSF,radarRE,radarMH) #fait tableau avec ensemble dimensions

labels = c("PF \n p=0.06","RP\n p<0.0001","BP\n p=0.25","HG\n p=0.18",
           "VT\n p=0.01","SF\n  p=0.0004","RE\n p=0.0008","MH\n p=0.33")

colnames(data)<-labels #remplace nom des colonnes par valeurs dans data

# Les données sont mises en forme sous forme de tableau dont les colonnes sont nommées
#requiert package fmsb et scales
library(fmsb)
library(scales)

datx = as.data.frame(data) ; datx

coul <- c("red", "blue") #creer un vecteur de 3 couleurs

radarchart(datx, 
           axistype = 2, 
           seg = 5, #radachart sur data frame datx, axistype= style des axes, seg=nb de graduation
           vlcex = 0.7,
           pcol = coul,#couleurs contour des polygones
           plwd=2 , 
           plty=1, #épaisseur lignes, et lignes pleines (plty=1) 
           cglcol="gray71", 
           cglty=1,
           axislabcol="grey",      #couleur et style grille

)


legend(1.6,0.8,#position légende
       legend=c("3 month","12 month"), 
       pch=c(16,16), #symboles points pleins
       col=c("red","blue"), #couleurs associées aux légendes
       lty=c(1,1), #ligne pleine
       text.font=1, #texte en gras
       box.lty=0 #pas de cadre autour figure
)
dev.off()
```


## Export data
```{r}
write.csv(bazR, "results/bazR_scores.csv", row.names = FALSE)
```

