---
title: "crussol.clementine"
author: "Clémentine Crussol"
output:
  html_document:
    df_print: paged
    code_folding: hide
    toc: false
    number_sections: true
---
# Impaired quality of life after intensive care

# 1. Scientific problem 
## 1.1. Context

Advances in intensive care medicine have significantly reduced ICU mortality over recent decades, leading to a growing population of ICU survivors with persistent physical, psychological, and cognitive impairments after discharge (1, 16). These long-term consequences, commonly grouped under the term Post-Intensive Care Syndrome (PICS), may substantially alter patients’ health-related quality of life (HRQoL) (13).

According to the World Health Organization (WHO), quality of life refers to an individual’s perception of their position in life within the cultural and social context in which they live (8). In the medical field, HRQoL is considered a multidimensional concept reflecting the influence of health status and healthcare on physical, psychological, and social dimensions of health (9). Previous studies have shown that ICU survivors often report lower HRQoL scores than the general population, with only partial recovery over time (18, 19).

This project aims to describe the evolution of HRQoL at 3 and 12 months after ICU discharge using the SF-36 questionnaire.

## 1.2. Aim

The aim of this study is to evaluate the quality of life of ICU survivors at 3 and 12 months after discharge, using the SF-36 questionnaire. Specifically, it seeks to assess changes over time in the different dimensions of quality of life, including physical and mental components, in order to better understand recovery trajectories following critical illness.

## 1.3. Method

All ICU survivors were contacted by mail 3 and 12 month after hospital discharge and were asked to complete the SF-36 questionnaire.

### Outcome measure

Quality of life was assesed using the SF-36 questionnaire. This questionnaire evaluate multiple dimensions of quality of life including : Physical Functionning (PF), Role Physical (RP), Bodily Pain (BP), General Health (GH), Vitality (VT), Social Functioning (SF), Role Emotional (RE), Mental Health (MH). 
Then, this dimensions are grouped into two components, the Physical Component Summary (PCS), including PF, RP, BP and GH, and the Mental Component Summary (MCS), including VT, SF, RE and MH.

SF-36 domains were scored according to standard procedures. Scores of each dimension were first computed by summing the values of all items belonging to that dimension and were then linearly transformed to a 0–100 scale. Higher scores indicate better heath status.

The Physical Component Summary (PCS) and Mental Component Summary (MCS) were then calculated using the dimensions scores. Then they were standardized unsing French reference population norms, with a mean of 50 and a standard deviation of 10.

Detailed information on SF-36 scoring procedures, including item mapping, score transformation, and computation of PCS and MCS, is provided in the `SF-36 appendix.md`.

### Participants

479 patients who survived ICU and responded to the SF-36 at both 3 and 12 months were included in the study.

# 2. RStudio project
## 2.1. Aim

The aim of this project is to assess SF-36 dimension scores at 3 and 12 months after ICU discharge. In addition, a radar chart will be generated to visualize changes in quality of life dimensions over time.

## 2.2. Data organization
File: `bazRSF.csv`  
Colums:  
`ID`: Patient ID, corresponding to their number of inclusion  
`SF_3m_Q[y]`: SF-36 score at 3 months for question y (where y is the question number e.g. 1, 2, 3a, ...)   
`SF_12m_Q[y]`: SF-36 score at 12 months for question y (where y is the question number e.g. 1, 2, 3a, ...)  

## 2.3. Script organization
### 2.3.1. Recoding of SF-36 item scores
**Aim**: To recode the SF-36 item responses so that higher scores correspond to a better health status accross all items.  
**Input**: `bazRSF.csv` in `data` folder  
**Output**: A set of SF-36 item scores in which higher values indicate better perceived health status.

### 2.3.2. Calculation and transformation of dimension scores at 3 and 12 months
**Aim**: To compute dimension scores and transform them onto a 0-100 scale for comparability and interpretation  
**Input**: `bazRSF.csv` in `data` folder  
**Calculation**:   
Raw scores for each dimension are first computed by summing the values of all items belonging to that dimension.  
Scores are then linearly transformed to a 0–100 scale using the following formula:  

$$
\text{Transformed score} = \frac{\text{Observed score} - \text{Minimum possible score}}{\text{Maximum possible score} - \text{Minimum possible score}} \times 100
$$

In this transformed scale, 0 represents the worst possible health status and 100 represents the best possible health status.  
**Output**: A standardized score for each dimension, ranging between 0 and 100  

### 2.3.3. Visualization of the evolution of dimension scores between 3 and 12 months
**Aim**: To visually compare dimension scores between 3 and 12 months, and assess if changes over time is statistically significant  
**Input**: `bazRSF.csv` in `data` folder  
**Output**: A radar plot comparing dimension scores at 3 and 12 months with p-values indicating whether changes over time reflect a significant improvement or not  




##  Import the data

```{r}
bazR <- read.csv("data/bazRSF.csv", header = TRUE, sep = ";")

```



```{r}
# Recode items for 3 months
bazR$SF_3m_Q1[bazR$SF_3m_Q1==1]<-5
bazR$SF_3m_Q1[bazR$SF_3m_Q1==2]<-4
bazR$SF_3m_Q1[bazR$SF_3m_Q1==3]<-3
bazR$SF_3m_Q1[bazR$SF_3m_Q1==4]<-2
bazR$SF_3m_Q1[bazR$SF_3m_Q1==5]<-1

   
bazR$SF_3m_Q2[bazR$SF_3m_Q2==1]<-5
bazR$SF_3m_Q2[bazR$SF_3m_Q2==2]<-4
bazR$SF_3m_Q2[bazR$SF_3m_Q2==3]<-3
bazR$SF_3m_Q2[bazR$SF_3m_Q2==4]<-2
bazR$SF_3m_Q2[bazR$SF_3m_Q2==5]<-1


bazR$SF_3m_Q6[bazR$SF_3m_Q6==1]<-5
bazR$SF_3m_Q6[bazR$SF_3m_Q6==2]<-4
bazR$SF_3m_Q6[bazR$SF_3m_Q6==3]<-3
bazR$SF_3m_Q6[bazR$SF_3m_Q6==4]<-2
bazR$SF_3m_Q6[bazR$SF_3m_Q6==5]<-1

 
bazR$SF_3m_Q7[bazR$SF_3m_Q7==1]<-6
bazR$SF_3m_Q7[bazR$SF_3m_Q7==2]<-5
bazR$SF_3m_Q7[bazR$SF_3m_Q7==3]<-4
bazR$SF_3m_Q7[bazR$SF_3m_Q7==4]<-3
bazR$SF_3m_Q7[bazR$SF_3m_Q7==5]<-2
bazR$SF_3m_Q7[bazR$SF_3m_Q7==6]<-1


bazR$SF_3m_Q8[bazR$SF_3m_Q8==1]<-5
bazR$SF_3m_Q8[bazR$SF_3m_Q8==2]<-4
bazR$SF_3m_Q8[bazR$SF_3m_Q8==3]<-3
bazR$SF_3m_Q8[bazR$SF_3m_Q8==4]<-2
bazR$SF_3m_Q8[bazR$SF_3m_Q8==5]<-1
 

bazR$SF_3m_Q9a[bazR$SF_3m_Q9a==1]<-6
bazR$SF_3m_Q9a[bazR$SF_3m_Q9a==2]<-5
bazR$SF_3m_Q9a[bazR$SF_3m_Q9a==3]<-4
bazR$SF_3m_Q9a[bazR$SF_3m_Q9a==4]<-3
bazR$SF_3m_Q9a[bazR$SF_3m_Q9a==5]<-2
bazR$SF_3m_Q9a[bazR$SF_3m_Q9a==6]<-1


bazR$SF_3m_Q9d[bazR$SF_3m_Q9d==1]<-6
bazR$SF_3m_Q9d[bazR$SF_3m_Q9d==2]<-5
bazR$SF_3m_Q9d[bazR$SF_3m_Q9d==3]<-4
bazR$SF_3m_Q9d[bazR$SF_3m_Q9d==4]<-3
bazR$SF_3m_Q9d[bazR$SF_3m_Q9d==5]<-2
bazR$SF_3m_Q9d[bazR$SF_3m_Q9d==6]<-1


bazR$SF_3m_Q9e[bazR$SF_3m_Q9e==1]<-6
bazR$SF_3m_Q9e[bazR$SF_3m_Q9e==2]<-5
bazR$SF_3m_Q9e[bazR$SF_3m_Q9e==3]<-4
bazR$SF_3m_Q9e[bazR$SF_3m_Q9e==4]<-5
bazR$SF_3m_Q9e[bazR$SF_3m_Q9e==5]<-2
bazR$SF_3m_Q9e[bazR$SF_3m_Q9e==6]<-1


bazR$SF_3m_Q9h[bazR$SF_3m_Q9h==1]<-6
bazR$SF_3m_Q9h[bazR$SF_3m_Q9h==2]<-5
bazR$SF_3m_Q9h[bazR$SF_3m_Q9h==3]<-4
bazR$SF_3m_Q9h[bazR$SF_3m_Q9h==4]<-3
bazR$SF_3m_Q9h[bazR$SF_3m_Q9h==5]<-2
bazR$SF_3m_Q9h[bazR$SF_3m_Q9h==6]<-1

 
bazR$SF_3m_Q11b[bazR$SF_3m_Q11b==1]<-5
bazR$SF_3m_Q11b[bazR$SF_3m_Q11b==2]<-4
bazR$SF_3m_Q11b[bazR$SF_3m_Q11b==3]<-3
bazR$SF_3m_Q11b[bazR$SF_3m_Q11b==4]<-2
bazR$SF_3m_Q11b[bazR$SF_3m_Q11b==5]<-1

 
bazR$SF_3m_Q11d[bazR$SF_3m_Q11d==1]<-5
bazR$SF_3m_Q11d[bazR$SF_3m_Q11d==2]<-4
bazR$SF_3m_Q11d[bazR$SF_3m_Q11d==3]<-3
bazR$SF_3m_Q11d[bazR$SF_3m_Q11d==4]<-2
bazR$SF_3m_Q11d[bazR$SF_3m_Q11d==5]<-1


# Recode items for 12 months

bazR$SF_12m_Q1[bazR$SF_12m_Q1==1]<-5
bazR$SF_12m_Q1[bazR$SF_12m_Q1==2]<-4
bazR$SF_12m_Q1[bazR$SF_12m_Q1==3]<-3
bazR$SF_12m_Q1[bazR$SF_12m_Q1==4]<-2
bazR$SF_12m_Q1[bazR$SF_12m_Q1==5]<-1

   
bazR$SF_12m_Q2[bazR$SF_12m_Q2==1]<-5
bazR$SF_12m_Q2[bazR$SF_12m_Q2==2]<-4
bazR$SF_12m_Q2[bazR$SF_12m_Q2==3]<-3
bazR$SF_12m_Q2[bazR$SF_12m_Q2==4]<-2
bazR$SF_12m_Q2[bazR$SF_12m_Q2==5]<-1


bazR$SF_12m_Q6[bazR$SF_12m_Q6==1]<-5
bazR$SF_12m_Q6[bazR$SF_12m_Q6==2]<-4
bazR$SF_12m_Q6[bazR$SF_12m_Q6==3]<-3
bazR$SF_12m_Q6[bazR$SF_12m_Q6==4]<-2
bazR$SF_12m_Q6[bazR$SF_12m_Q6==5]<-1

 
bazR$SF_12m_Q7[bazR$SF_12m_Q7==1]<-6
bazR$SF_12m_Q7[bazR$SF_12m_Q7==2]<-5
bazR$SF_12m_Q7[bazR$SF_12m_Q7==3]<-4
bazR$SF_12m_Q7[bazR$SF_12m_Q7==4]<-3
bazR$SF_12m_Q7[bazR$SF_12m_Q7==5]<-2
bazR$SF_12m_Q7[bazR$SF_12m_Q7==6]<-1


bazR$SF_12m_Q8[bazR$SF_12m_Q8==1]<-5
bazR$SF_12m_Q8[bazR$SF_12m_Q8==2]<-4
bazR$SF_12m_Q8[bazR$SF_12m_Q8==3]<-3
bazR$SF_12m_Q8[bazR$SF_12m_Q8==4]<-2
bazR$SF_12m_Q8[bazR$SF_12m_Q8==5]<-1
 

bazR$SF_12m_Q9a[bazR$SF_12m_Q9a==1]<-6
bazR$SF_12m_Q9a[bazR$SF_12m_Q9a==2]<-5
bazR$SF_12m_Q9a[bazR$SF_12m_Q9a==3]<-4
bazR$SF_12m_Q9a[bazR$SF_12m_Q9a==4]<-3
bazR$SF_12m_Q9a[bazR$SF_12m_Q9a==5]<-2
bazR$SF_12m_Q9a[bazR$SF_12m_Q9a==6]<-1


bazR$SF_12m_Q9d[bazR$SF_12m_Q9d==1]<-6
bazR$SF_12m_Q9d[bazR$SF_12m_Q9d==2]<-5
bazR$SF_12m_Q9d[bazR$SF_12m_Q9d==3]<-4
bazR$SF_12m_Q9d[bazR$SF_12m_Q9d==4]<-3
bazR$SF_12m_Q9d[bazR$SF_12m_Q9d==5]<-2
bazR$SF_12m_Q9d[bazR$SF_12m_Q9d==6]<-1


bazR$SF_12m_Q9e[bazR$SF_12m_Q9e==1]<-6
bazR$SF_12m_Q9e[bazR$SF_12m_Q9e==2]<-5
bazR$SF_12m_Q9e[bazR$SF_12m_Q9e==3]<-4
bazR$SF_12m_Q9e[bazR$SF_12m_Q9e==4]<-5
bazR$SF_12m_Q9e[bazR$SF_12m_Q9e==5]<-2
bazR$SF_12m_Q9e[bazR$SF_12m_Q9e==6]<-1


bazR$SF_12m_Q9h[bazR$SF_12m_Q9h==1]<-6
bazR$SF_12m_Q9h[bazR$SF_12m_Q9h==2]<-5
bazR$SF_12m_Q9h[bazR$SF_12m_Q9h==3]<-4
bazR$SF_12m_Q9h[bazR$SF_12m_Q9h==4]<-3
bazR$SF_12m_Q9h[bazR$SF_12m_Q9h==5]<-2
bazR$SF_12m_Q9h[bazR$SF_12m_Q9h==6]<-1

 
bazR$SF_12m_Q11b[bazR$SF_12m_Q11b==1]<-5
bazR$SF_12m_Q11b[bazR$SF_12m_Q11b==2]<-4
bazR$SF_12m_Q11b[bazR$SF_12m_Q11b==3]<-3
bazR$SF_12m_Q11b[bazR$SF_12m_Q11b==4]<-2
bazR$SF_12m_Q11b[bazR$SF_12m_Q11b==5]<-1

 
bazR$SF_12m_Q11d[bazR$SF_12m_Q11d==1]<-5
bazR$SF_12m_Q11d[bazR$SF_12m_Q11d==2]<-4
bazR$SF_12m_Q11d[bazR$SF_12m_Q11d==3]<-3
bazR$SF_12m_Q11d[bazR$SF_12m_Q11d==4]<-2
bazR$SF_12m_Q11d[bazR$SF_12m_Q11d==5]<-1
```

## Calculation of dimension scores at 3 months
### Imputation of missing data

```{r}
items_sf36 <- list(
  PF = c("SF_3m_Q3a", "SF_3m_Q3b","SF_3m_Q3c","SF_3m_Q3d", "SF_3m_Q3e", "SF_3m_Q3f","SF_3m_Q3g",
         "SF_3m_Q3h","SF_3m_Q3i","SF_3m_Q3j"),
  RP = c("SF_3m_Q4a", "SF_3m_Q4b","SF_3m_Q4c","SF_3m_Q4d"),
  BP = c("SF_3m_Q7","SF_3m_Q8"),
  GH = c("SF_3m_Q1","SF_3m_Q11a","SF_3m_Q11b", "SF_3m_Q11c","SF_3m_Q11d"),
  VT = c("SF_3m_Q9a", "SF_3m_Q9e","SF_3m_Q9g","SF_3m_Q9i"),
  SF = c("SF_3m_Q6", "SF_3m_Q10"),
  RE = c("SF_3m_Q5a","SF_3m_Q5b","SF_3m_Q5c"),
  MH = c("SF_3m_Q9b","SF_3m_Q9c","SF_3m_Q9d","SF_3m_Q9f","SF_3m_Q9h")
)

impute_sf36 <- function(df, items_list) {
  df_out <- df
  
  for (dim in names(items_list)) {
    vars <- items_list[[dim]]
    
    # nombre d'items présents par patient
    present <- rowSums(!is.na(df[vars]))
    
    # règle SF-36 : au moins 50% des items présents
    enough <- present >= length(vars) / 2
    
    # moyenne des items présents
    row_means <- rowMeans(df[vars], na.rm = TRUE)
    
    # imputation des NA si règle respectée
    for (v in vars) {
      df_out[[v]][is.na(df[[v]]) & enough] <- row_means[is.na(df[[v]]) & enough]
    }
    
    # si < 50% → toute la dimension devient NA
    df_out[!enough, vars] <- NA
  }
  
  return(df_out)
}
bazR <- impute_sf36(bazR, items_sf36)
```

### Calculation and transformation of dimension scores at 3 months
Raw scores for each dimension were first computed by summing the values of all items belonging to that dimension.
Scores were then linearly transformed to a 0–100 scale using the following formula:

Transformed score = ((Observed score - Minimum possible score) / (Maximum possible score - Minimum possible score)) × 100

In this transformed scale, 0 represents the worst possible health status and 100 represents the best possible health status.

```{r}
# Dimension PF
aPF_3m <- (bazR$SF_3m_Q3a + bazR$SF_3m_Q3b + bazR$SF_3m_Q3c + bazR$SF_3m_Q3d + bazR$SF_3m_Q3e + bazR$SF_3m_Q3f
            + bazR$SF_3m_Q3g + bazR$SF_3m_Q3h + bazR$SF_3m_Q3i + bazR$SF_3m_Q3j)

bazR$PFtransformed_3m <- ((aPF_3m-10)/20)*100 


# Dimension RP  
aRP_3m <- (bazR$SF_3m_Q4a + bazR$SF_3m_Q4b + bazR$SF_3m_Q4c + bazR$SF_3m_Q4d)

bazR$RPtransformed_3m<-((aRP_3m-4)/4)*100

# Dimension BP
aBP_3m <- (bazR$SF_3m_Q7 + bazR$SF_3m_Q8)

bazR$BPtransformed_3m <- ((aBP_3m-2)/10)*100

# Dimension GH
aGH_3m <- (bazR$SF_3m_Q1 + bazR$SF_3m_Q11a + bazR$SF_3m_Q11b + bazR$SF_3m_Q11c + bazR$SF_3m_Q11d)

bazR$GHtransformed_3m <- ((aGH_3m-5)/20)*100

# Dimension VT
aVT_3m <- (bazR$SF_3m_Q9a + bazR$SF_3m_Q9e + bazR$SF_3m_Q9g + bazR$SF_3m_Q9i)

bazR$VTtransformed_3m <- ((aVT_3m-4)/20)*100

# Dimension SF
aSF_3m <- (bazR$SF_3m_Q6 + bazR$SF_3m_Q10)

bazR$SFtransformed_3m <- ((aSF_3m-2)/8)*100

#calcul dimension RE
aRE_3m <- (bazR$SF_3m_Q5a + bazR$SF_3m_Q5b + bazR$SF_3m_Q5c)

bazR$REtransformed_3m <- ((aRE_3m-3)/3)*100

# Dimension MH
aMH_3m <- (bazR$SF_3m_Q9b + bazR$SF_3m_Q9c + bazR$SF_3m_Q9d + bazR$SF_3m_Q9f + bazR$SF_3m_Q9h)

bazR$MHtransformed_3m <- ((aMH_3m-5)/25)*100

# Dimension HT
aHT_3m <- bazR$SF_3m_Q2

bazR$HTtransformed_3m <- ((aHT_3m-1)/4)*100

```


## Calculation of dimension scores at 12 months
### Imputation of missing data

```{r}

items_sf36 <- list(
  PF = c("SF_12m_Q3a", "SF_12m_Q3b","SF_12m_Q3c","SF_12m_Q3d", "SF_12m_Q3e", "SF_12m_Q3f",
         "SF_12m_Q3g", "SF_12m_Q3h","SF_12m_Q3i","SF_12m_Q3j"),
  RP = c("SF_12m_Q4a", "SF_12m_Q4b", "SF_12m_Q4c","SF_12m_Q4d"),
  BP = c("SF_12m_Q7","SF_12m_Q8"),
  GH = c("SF_12m_Q1","SF_12m_Q11a","SF_12m_Q11b", "SF_12m_Q11c","SF_12m_Q11d"),
  VT = c("SF_12m_Q9a", "SF_12m_Q9e","SF_12m_Q9g","SF_12m_Q9i"),
  SF = c("SF_12m_Q6", "SF_12m_Q10"),
  RE = c("SF_12m_Q5a","SF_12m_Q5b","SF_12m_Q5c"),
  MH = c("SF_12m_Q9b","SF_12m_Q9c","SF_12m_Q9d","SF_12m_Q9f","SF_12m_Q9h")
)

impute_sf36 <- function(df, items_list) {
  df_out <- df
  
  for (dim in names(items_list)) {
    vars <- items_list[[dim]]
    
    # nombre d'items présents par patient
    present <- rowSums(!is.na(df[vars]))
    
    # règle SF-36 : au moins 50% des items présents
    enough <- present >= length(vars) / 2
    
    # moyenne des items présents
    row_means <- rowMeans(df[vars], na.rm = TRUE)
    
    # imputation des NA si règle respectée
    for (v in vars) {
      df_out[[v]][is.na(df[[v]]) & enough] <- row_means[is.na(df[[v]]) & enough]
    }
    
    # si < 50% → toute la dimension devient NA
    df_out[!enough, vars] <- NA
  }
  
  return(df_out)
  
}
bazR <- impute_sf36(bazR, items_sf36)

```

### Calculation and transformation of dimension scores at 3 months
Raw scores for each dimension were first computed by summing the values of all items belonging to that dimension.
Scores were then linearly transformed to a 0–100 scale using the following formula:

Transformed score = ((Observed score - Minimum possible score) / (Maximum possible score - Minimum possible score)) × 100

In this transformed scale, 0 represents the worst possible health status and 100 represents the best possible health status.

```{r}
# Dimension PF
aPF_12m <- (bazR$SF_12m_Q3a + bazR$SF_12m_Q3b + bazR$SF_12m_Q3c + bazR$SF_12m_Q3d + bazR$SF_12m_Q3e 
            + bazR$SF_12m_Q3f + bazR$SF_12m_Q3g + bazR$SF_12m_Q3h + bazR$SF_12m_Q3i + bazR$SF_12m_Q3j)

bazR$PFtransformed_12m<-((aPF_12m-10)/20)*100 

# Dimension RP
aRP_12m <- (bazR$SF_12m_Q4a + bazR$SF_12m_Q4b + bazR$SF_12m_Q4c + bazR$SF_12m_Q4d)

bazR$RPtransformed_12m <- ((aRP_12m-4)/4)*100

# Dimension BP
aBP_12m <- (bazR$SF_12m_Q7 + bazR$SF_12m_Q8)

bazR$BPtransformed_12m <- ((aBP_12m-2)/10)*100

# Dimension GH
aGH_12m <- (bazR$SF_12m_Q1 + bazR$SF_12m_Q11a + bazR$SF_12m_Q11b + bazR$SF_12m_Q11c + bazR$SF_12m_Q11d)

bazR$GHtransformed_12m <- ((aGH_12m-5)/20)*100

# Dimension VT
aVT_12m <- (bazR$SF_12m_Q9a + bazR$SF_12m_Q9e + bazR$SF_12m_Q9g + bazR$SF_12m_Q9i)

bazR$VTtransformed_12m <- ((aVT_12m-4)/20)*100

# Dimension SF
aSF_12m <- (bazR$SF_12m_Q6 + bazR$SF_12m_Q10)

bazR$SFtransformed_12m <- ((aSF_12m-2)/8)*100

# Dimension RE
aRE_12m <- (bazR$SF_12m_Q5a + bazR$SF_12m_Q5b + bazR$SF_12m_Q5c)

bazR$REtransformed_12m <- ((aRE_12m-3)/3)*100

# Dimension MH
aMH_12m <- (bazR$SF_12m_Q9b + bazR$SF_12m_Q9c + bazR$SF_12m_Q9d + bazR$SF_12m_Q9f + bazR$SF_12m_Q9h)

bazR$MHtransformed_12m <- ((aMH_12m-5)/25)*100

# Dimension HT
aHT_12m <- bazR$SF_12m_Q2

bazR$HTtransformed_12m <- ((aHT_12m-1)/4)*100

```

## Statistical comparison of dimension scores at 3 and 12 months and construction of a radar plot

```{r}

summary(bazR$PFtransformed_3m)
summary(bazR$PFtransformed_12m)
radarPF=c(100,0,57.94,61.19) #max 100, min 0, PF3mois=58.19, PF1an=62.27
wilcox.test(bazR$PFtransformed_3m,bazR$PFtransformed_12m)


summary(bazR$RPtransformed_3m)
summary(bazR$RPtransformed_12m)
radarRP=c(100,0,31.6,47.03)
wilcox.test(bazR$RPtransformed_3m,bazR$RPtransformed_12m)


summary(bazR$BPtransformed_3m)
summary(bazR$BPtransformed_12m)
radarBP=c(100,0,46.64,48.16)
wilcox.test(bazR$BPtransformed_3m,bazR$BPtransformed_12m)


summary(bazR$GHtransformed_3m)
summary(bazR$GHtransformed_12m)
radarHG=c(100,0,50.42,52.3)
wilcox.test(bazR$GHtransformed_3m,bazR$GHtransformed_12m)


summary(bazR$VTtransformed_3m)
summary(bazR$VTtransformed_12m)
radarVT=c(100,0,47.53,51.22)
wilcox.test(bazR$VTtransformed_3m,bazR$VTtransformed_12m)


summary(bazR$SFtransformed_3m)
summary(bazR$SFtransformed_12m)
radarSF=c(100,0,61.01,67.09)
wilcox.test(bazR$SFtransformed_3m,bazR$SFtransformed_12m)


summary(bazR$REtransformed_3m)
summary(bazR$REtransformed_12m)
radarRE=c(100,0,46.19,56.06)
wilcox.test(bazR$REtransformed_3m,bazR$REtransformed_12m)


summary(bazR$MHtransformed_3m)
summary(bazR$MHtransformed_12m)
radarMH=c(100,0,61.35,62.45)
wilcox.test(bazR$MHtransformed_3m,bazR$MHtransformed_12m)

png("./results/radar_plot.png",width = 1800,height = 1800,res = 300)

data<-cbind(radarPF,radarRP,radarBP,radarHG,radarVT,radarSF,radarRE,radarMH) #fait tableau avec ensemble dimensions

labels = c("PF \n p=0.06","RP\n p<0.0001","BP\n p=0.25","HG\n p=0.18",
           "VT\n p=0.01","SF\n  p=0.0004","RE\n p=0.0008","MH\n p=0.33")

colnames(data)<-labels #remplace nom des colonnes par valeurs dans data

# Les données sont mises en forme sous forme de tableau dont les colonnes sont nommées
#requiert package fmsb et scales
library(fmsb)
library(scales)

datx = as.data.frame(data) ; datx

coul <- c("red", "blue") #creer un vecteur de 3 couleurs

radarchart(datx, 
           axistype = 2, 
           seg = 5, #radachart sur data frame datx, axistype= style des axes, seg=nb de graduation
           vlcex = 0.7,
           pcol = coul,#couleurs contour des polygones
           plwd=2 , 
           plty=1, #épaisseur lignes, et lignes pleines (plty=1) 
           cglcol="gray71", 
           cglty=1,
           axislabcol="grey",      #couleur et style grille

)


legend(1.6,0.8,#position légende
       legend=c("3 month","12 month"), 
       pch=c(16,16), #symboles points pleins
       col=c("red","blue"), #couleurs associées aux légendes
       lty=c(1,1), #ligne pleine
       text.font=1, #texte en gras
       box.lty=0 #pas de cadre autour figure
)
dev.off()
```


## Export data
```{r}
write.csv(bazR, "results/bazP_scores.csv", row.names = FALSE)
```

